---
title: "Laboritory session 1"
author: "Benedikt Schneider, Carolina Dobras"
date: "\\today"

header-includes:
  - \usepackage{csquotes}
  - \usepackage{titling}
  - \setlength{\droptitle}{-3cm}
output:
  pdf_document:
    
    highlight: tango
    keep_tex: no
    latex_engine: xelatex
    number_sections: yes
    toc: no
    toc_depth: 2
  html_document:
    
    theme: default
    df_print: paged
    toc: no
    toc_depth: '2'
    number_sections: yes
    highlight: tango
    toc_float: yes
    code_folding: hide
link_citation: yes
mainfont: XCharter
lang: de-DE
---

```{r setup, include=FALSE}
library(knitr)
library(Rcpp)
library(reticulate)
use_python("/usr/bin/python3")
knitr::opts_chunk$set(fig.align='center',
                      fig.width=6.0,
                      fig.height=3.6)
```
```{r, include = FALSE}
Sys.setenv("PKG_CXXFLAGS"="-std=c++17")
library(Rcpp)
```

# Laboratory session 1 

The first experiment is using standing waves inside a long tube in order to measure the speed of sound inside the gas the tube is filled with. The tube and 
experiment settings are specified by the following data: 

Tube length: $L = 1244 \pm 1.5mm$. 

input signal: $V_0 = 35mV$

Tubes: 

K1: argon or $CO_2$ at $T = T_{room}$

K2: air at $T = T_{room}$

K3: air at $T = 70\,^\circ\mathrm{C}$ = $343.15K$

K4: air at $T = 50\,^\circ\mathrm{C}$ = $323.15$

$T_{room} = 21\,^\circ\mathrm{C}$ = $294.15$

We analyzed the properties of the gas in tube K4. The resistance of the tube is measured to be $R = (34.55 \pm 0.01) k\Omega$.

# Tasks

## Meaurement of $\nu _k$

We measured the following resonance frequencies $\nu _k$ between $200 Hz$ and $2 kHz$:

K4: $\nu _k = [313 \pm 1, 450 \pm 1, 594 \pm 1, 740 \pm 1, 886 \pm 1, 1033 \pm 1, 1178 \pm 1, 1326 \pm 1, 1473 \pm 1, 1620 \pm 1, 1767 \pm 1, 1913 \pm 1]$

The other groups meaured:

K3: $\nu _k = [301 \pm 1, 459 \pm 1, 607 \pm 1, 757 \pm 1, 907 \pm 1, 1057 \pm 1, 1208 \pm 1, 1359 \pm 1, 1510 \pm 1, 1661 \pm 1, 1813 \pm 1, 1963 \pm 1]$

K2: $\nu _k = [286 \pm 1, 419 \pm 1, 470 \pm 1, 563 \pm 1, 700 \pm 1, 840 \pm 1,979 \pm 1,1117 \pm 1,1258 \pm 1,1397 \pm 1,1537 \pm 1,1677 \pm 1,1816 \pm 1,1955 \pm 1]$

K1: argon: $\nu _k = [265 \pm 1, 393 \pm 1, 521 \pm 1, 650 \pm 1, 778\pm 1, 905 \pm 1, 1035 \pm 1, 1164 \pm 1, 1293\pm 1, 1423 \pm 1, 1551 \pm 1, 1680 \pm 1, 1810 \pm 1, 1938\pm 1]$

$CO_2$: $\nu _k = [225 \pm 1, 440 \pm 1, 544 \pm 1, 656 \pm 1, 764\pm 1, 872 \pm 1, 979 \pm 1, 1091\pm 1, 1199 \pm 1, 1209 \pm 1, 1418 \pm 1, 1527 \pm 1, 1635 \pm 1, 1745 \pm 1, 1854\pm 1, 1963\pm 1]$

## Estimated error 

We estimated our error to be around $1Hz$

## The best linear fit 

In order to find the speed of sound in the gas, we want to use the equation $\nu _n = an + b$. However, we do not know which resonance ferquency we are
measuring, so we do not know which n belongs to which measured $\nu _n$. This is why we set the reference of the measured $\nu _n$ to the first one measured. Therefore we get:

$$\nu _n - \nu _{n_0} = \Delta \nu _{\Delta n} = a\Delta n + b - (a(n_0) + b) = a(n-n_0) + b - b = a\Delta n = \frac{c}{2L} \Delta n$$

So the central relation is: $$\Delta \nu _{\Delta n} = \frac{c}{2L} \Delta n$$ 

where $\Delta n$ starts at 0 for the first measurement and for $k$ measurements continues until $k-1$ for the $k^{th}$ resonance frequency. $\Delta \nu _{\Delta n}$ against $\Delta n$ is plotted and
evaluated in the following:

K4:

```{python, echo = FALSE}
from uncertainties import unumpy as u, covariance_matrix as co, ufloat
import numpy as np
from scipy.optimize import curve_fit
import matplotlib.pyplot as plt

nu = u.uarray([313,450,594,740,886,1033,1178,1326,1473,1620,1767,1913], [1,1,1,1,1,1,1,1,1,1,1,1]) #[Messwerte], [sigmas] !
d_n = np.arange(len(nu))
dnu = u.uarray([],[])

for i in range(len(nu)):
    l = nu[i] - nu[0]
    dnu = np.append(dnu, l)

dnu[0] = ufloat(u.nominal_values(dnu)[0],np.mean(u.std_devs(dnu))) #set std dev of the first difference to the mean std dev because otherwise it would be 0

def lin(x, a, b):
    return a*x +b

popt, pcov = curve_fit(lin, d_n, u.nominal_values(dnu), sigma = co(dnu), absolute_sigma = True) 

a = u.uarray([popt[0]], [np.sqrt(pcov[0,0])])
b = popt[1]

plt.figure()
plt.errorbar(d_n,  u.nominal_values(dnu), yerr = u.std_devs(dnu), ecolor = "darkred", capsize = 5, label = "Datapoints", color = "darkblue", fmt = "o")
plt.plot(d_n, u.nominal_values(a)*d_n + b, label = f"Fitted curve, slope = {u.nominal_values(a)[0]:.2f} +- {u.std_devs(a)[0]:.2f}", color = "darkgreen")
plt.grid()
plt.legend()
plt.xlabel(r"$\Delta n$")
plt.ylabel(r"$\Delta \nu _{\Delta n} [Hz]$")
plt.show()
```

Given the slope of the fit curve, the velocity of sound can be calculated:

```{python, echo = FALSE, comment = NA}
L = ufloat(1.244,0.0015)  #Länge, Fehler auf Länge
v_sound_k4 = 2*L*a
print(f"speed of sound = {u.nominal_values(v_sound_k4)[0]:.2f} +- {u.std_devs(v_sound_k4)[0]:.2f} m/s")

```
K3:

The resistance of the tube is measured to be $R = (15.66 \pm 0.02) k\Omega$

```{python, echo = FALSE}
from uncertainties import unumpy as u, covariance_matrix as co, ufloat
import numpy as np
from scipy.optimize import curve_fit
import matplotlib.pyplot as plt

nu = u.uarray([301, 459, 607, 757, 907, 1057, 1208, 1359, 1510, 1661, 1813 , 1963], [1,1,1,1,1,1,1,1,1,1,1,1]) #[Messwerte], [sigmas] !
d_n = np.arange(len(nu))
dnu = u.uarray([],[])

for i in range(len(nu)):
    l = nu[i] - nu[0]
    dnu = np.append(dnu, l)

dnu[0] = ufloat(u.nominal_values(dnu)[0],np.mean(u.std_devs(dnu))) #set std dev of the first difference to the mean std dev because otherwise it would be 0

def lin(x, a, b):
    return a*x +b

popt, pcov = curve_fit(lin, d_n, u.nominal_values(dnu), sigma = co(dnu), absolute_sigma = True) 

a = u.uarray([popt[0]], [np.sqrt(pcov[0,0])])
b = popt[1]

plt.figure()
plt.errorbar(d_n,  u.nominal_values(dnu), yerr = u.std_devs(dnu), ecolor = "darkred", capsize = 5, label = "Datapoints", color = "darkblue", fmt = "o")
plt.plot(d_n, u.nominal_values(a)*d_n + b, label = f"Fitted curve, slope = {u.nominal_values(a)[0]:.2f} +- {u.std_devs(a)[0]:.2f}", color = "darkgreen")
plt.grid()
plt.legend()
plt.xlabel(r"$\Delta n$")
plt.ylabel(r"$\Delta \nu _{\Delta n} [Hz]$")
plt.show()
```

Given the slope of the fit curve, the velocity of sound can be calculated:

```{python, echo = FALSE, comment = NA}
v_sound_k3 = 2*L*a
print(f"speed of sound = {u.nominal_values(v_sound_k3)[0]:.2f} +- {u.std_devs(v_sound_k3)[0]:.2f} m/s")

```
K2:

```{python, echo = FALSE}
from uncertainties import unumpy as u, covariance_matrix as co, ufloat
import numpy as np
from scipy.optimize import curve_fit
import matplotlib.pyplot as plt

nu = u.uarray([286, 419, 563 , 700, 840 ,979 ,1117 ,1258 ,1397 ,1537 ,1677 ,1816 ,1955], [1,1,1,1,1,1,1,1,1,1,1,1,1]) #[Messwerte], [sigmas] !
d_n = np.arange(len(nu))
dnu = u.uarray([],[])

for i in range(len(nu)):
    l = nu[i] - nu[0]
    dnu = np.append(dnu, l)

dnu[0] = ufloat(u.nominal_values(dnu)[0],np.mean(u.std_devs(dnu))) #set std dev of the first difference to the mean std dev because otherwise it would be 0

def lin(x, a, b):
    return a*x +b

popt, pcov = curve_fit(lin, d_n, u.nominal_values(dnu), sigma = co(dnu), absolute_sigma = True) 

a = u.uarray([popt[0]], [np.sqrt(pcov[0,0])])
b = popt[1]

plt.figure()
plt.errorbar(d_n,  u.nominal_values(dnu), yerr = u.std_devs(dnu), ecolor = "darkred", capsize = 5, label = "Datapoints", color = "darkblue", fmt = "o")
plt.plot(d_n, u.nominal_values(a)*d_n + b, label = f"Fitted curve, slope = {u.nominal_values(a)[0]:.2f} +- {u.std_devs(a)[0]:.2f}", color = "darkgreen")
plt.grid()
plt.legend()
plt.xlabel(r"$\Delta n$")
plt.ylabel(r"$\Delta \nu _{\Delta n} [Hz]$")
plt.show()
```

Given the slope of the fit curve, the velocity of sound can be calculated:

```{python, echo = FALSE, comment = NA}
v_sound_k2 = 2*L*a
print(f"speed of sound = {u.nominal_values(v_sound_k2)[0]:.2f} +- {u.std_devs(v_sound_k2)[0]:.2f} m/s")

```
K1, argon:

```{python, echo = FALSE}
from uncertainties import unumpy as u, covariance_matrix as co, ufloat
import numpy as np
from scipy.optimize import curve_fit
import matplotlib.pyplot as plt

nu = u.uarray([265 ,393 ,521 ,650 ,778, 905 ,1035 ,1164 ,1293, 1423 ,1551 ,1680 ,1810 ,1938], [1,1,1,1,1,1,1,1,1,1,1,1,1,1]) #[Messwerte], [sigmas] !
d_n = np.arange(len(nu))
dnu = u.uarray([],[])

for i in range(len(nu)):
    l = nu[i] - nu[0]
    dnu = np.append(dnu, l)

dnu[0] = ufloat(u.nominal_values(dnu)[0],np.mean(u.std_devs(dnu))) #set std dev of the first difference to the mean std dev because otherwise it would be 0

def lin(x, a, b):
    return a*x +b

popt, pcov = curve_fit(lin, d_n, u.nominal_values(dnu), sigma = co(dnu), absolute_sigma = True) 

a = u.uarray([popt[0]], [np.sqrt(pcov[0,0])])
b = popt[1]

plt.figure()
plt.errorbar(d_n,  u.nominal_values(dnu), yerr = u.std_devs(dnu), ecolor = "darkred", capsize = 5, label = "Datapoints", color = "darkblue", fmt = "o")
plt.plot(d_n, u.nominal_values(a)*d_n + b, label = f"Fitted curve, slope = {u.nominal_values(a)[0]:.2f} +- {u.std_devs(a)[0]:.2f}", color = "darkgreen")
plt.grid()
plt.legend()
plt.xlabel(r"$\Delta n$")
plt.ylabel(r"$\Delta \nu _{\Delta n} [Hz]$")
plt.show()
```

Given the slope of the fit curve, the velocity of sound can be calculated:

```{python, echo = FALSE, comment = NA}
v_sound_k1_argon = 2*L*a
print(f"speed of sound = {u.nominal_values(v_sound_k1_argon)[0]:.2f} +- {u.std_devs(v_sound_k1_argon)[0]:.2f} m/s")

```
K1, $CO_2$:

```{python, echo = FALSE}
from uncertainties import unumpy as u, covariance_matrix as co, ufloat
import numpy as np
from scipy.optimize import curve_fit
import matplotlib.pyplot as plt

nu = u.uarray([225,440,544,656,764,872,979,1091,1199,1209,1418,1527,1635,1745,1854,1963], [1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1]) #[Messwerte], [sigmas] !
d_n = np.arange(len(nu))
dnu = u.uarray([],[])

for i in range(len(nu)):
    l = nu[i] - nu[0]
    dnu = np.append(dnu, l)

dnu[0] = ufloat(u.nominal_values(dnu)[0],np.mean(u.std_devs(dnu))) #set std dev of the first difference to the mean std dev because otherwise it would be 0

def lin(x, a, b):
    return a*x +b

popt, pcov = curve_fit(lin, d_n, u.nominal_values(dnu), sigma = co(dnu), absolute_sigma = True) 

a = u.uarray([popt[0]], [np.sqrt(pcov[0,0])])
b = popt[1]

plt.figure()
plt.errorbar(d_n,  u.nominal_values(dnu), yerr = u.std_devs(dnu), ecolor = "darkred", capsize = 5, label = "Datapoints", color = "darkblue", fmt = "o")
plt.plot(d_n, u.nominal_values(a)*d_n + b, label = f"Fitted curve, slope = {u.nominal_values(a)[0]:.2f} +- {u.std_devs(a)[0]:.2f}", color = "darkgreen")
plt.grid()
plt.legend()
plt.xlabel(r"$\Delta n$")
plt.ylabel(r"$\Delta \nu _{\Delta n} [Hz]$")
plt.show()
```

Given the slope of the fit curve, the velocity of sound can be calculated:

```{python, echo = FALSE, comment = NA}
v_sound_k1_co2 = 2*L*a
print(f"speed of sound = {u.nominal_values(v_sound_k1_co2)[0]:.2f} +- {u.std_devs(v_sound_k1_co2)[0]:.2f} m/s")

```
## Most important contribution to uncertainty

The most important contribution to the error is the inaccuracy in finetuning the resonance frequency. We estimated our accuracy in doing so to $1Hz$.

The uncertainty shrinks the more points are included in the fit. In the following we see the plot for K4, once with 12 points and once with 6:


```{python, echo = FALSE}
from uncertainties import unumpy as u, covariance_matrix as co, ufloat
import numpy as np
from scipy.optimize import curve_fit
import matplotlib.pyplot as plt

nu = u.uarray([313,450,594,740,886,1033,1178,1326,1473,1620,1767,1913], [1,1,1,1,1,1,1,1,1,1,1,1]) #[Messwerte], [sigmas] !
d_n = np.arange(len(nu))
dnu = u.uarray([],[])

for i in range(len(nu)):
    l = nu[i] - nu[0]
    dnu = np.append(dnu, l)

dnu[0] = ufloat(u.nominal_values(dnu)[0],np.mean(u.std_devs(dnu))) #set std dev of the first difference to the mean std dev because otherwise it would be 0

def lin(x, a, b):
    return a*x +b

popt, pcov = curve_fit(lin, d_n, u.nominal_values(dnu), sigma = co(dnu), absolute_sigma = True) 

a = u.uarray([popt[0]], [np.sqrt(pcov[0,0])])
b = popt[1]

plt.figure()
plt.errorbar(d_n,  u.nominal_values(dnu), yerr = u.std_devs(dnu), ecolor = "darkred", capsize = 5, label = "Datapoints", color = "darkblue", fmt = "o")
plt.plot(d_n, u.nominal_values(a)*d_n + b, label = f"Fitted curve, slope = {u.nominal_values(a)[0]:.2f} +- {u.std_devs(a)[0]:.2f}", color = "darkgreen")
plt.grid()
plt.legend()
plt.title("K4, 12 points")
plt.xlabel(r"$\Delta n$")
plt.ylabel(r"$\Delta \nu _{\Delta n} [Hz]$")
plt.show()
```


```{python, echo = FALSE}
from uncertainties import unumpy as u, covariance_matrix as co, ufloat
import numpy as np
from scipy.optimize import curve_fit
import matplotlib.pyplot as plt

nu = u.uarray([313,594,886,1178,1473,1767], [1,1,1,1,1,1]) #[Messwerte], [sigmas] !
d_n = np.arange(len(nu))*2
dnu = u.uarray([],[])

for i in range(len(nu)):
    l = nu[i] - nu[0]
    dnu = np.append(dnu, l)

dnu[0] = ufloat(u.nominal_values(dnu)[0],np.mean(u.std_devs(dnu))) #set std dev of the first difference to the mean std dev because otherwise it would be 0

def lin(x, a, b):
    return a*x +b

popt, pcov = curve_fit(lin, d_n, u.nominal_values(dnu), sigma = co(dnu), absolute_sigma = True) 

a = u.uarray([popt[0]], [np.sqrt(pcov[0,0])])
b = popt[1]

plt.figure()
plt.errorbar(d_n,  u.nominal_values(dnu), yerr = u.std_devs(dnu), ecolor = "darkred", capsize = 5, label = "Datapoints", color = "darkblue", fmt = "o")
plt.plot(d_n, u.nominal_values(a)*d_n + b, label = f"Fitted curve, slope = {u.nominal_values(a)[0]:.2f} +- {u.std_devs(a)[0]:.2f}", color = "darkgreen")
plt.grid()
plt.legend()
plt.title("K4, 6 points")
plt.xlabel(r"$\Delta n$")
plt.ylabel(r"$\Delta \nu _{\Delta n} [Hz]$")
plt.show()
```

As we can see, the uncertainty on the slope is more than 50% greater if we only choose half of the points.

## Degrees of freedom

For an ideal gas, the velocity of sound is given by $$c(T) = \sqrt{\frac{(f+2)RT}{fM_{mol}}}$$ where $f$ are the degrees of freedom, $R$ is the gas constant,
$T$ the temperature of the gas and $M_mol$ the molar mass. This equation can be soved for $f$: $$f = \frac{2RT}{c^2M_{mol} - RT}$$
We calculate for $f$:

K4:

```{python,echo = FALSE, comment = NA}
R = 8.314462618
T = ufloat(343.15,1) #Temperature + error in [K]
c = v_sound_k4
M_mol = 0.02896 #[molare masse] [kg/mol]

f = 2*R*T/(c**2*M_mol - R*T)
print(f"The calculated amount of degrees of freedom from the experiment is {u.nominal_values(f)[0]:.2f} +- {u.std_devs(f)[0]:.2f}")
```
K3:

```{python,echo = FALSE, comment = NA}
R = 8.314462618
T = ufloat(323.15,1) #Temperature + error in [K]
c = v_sound_k3
M_mol = 0.02896 #[molare masse] [kg/mol]

f = 2*R*T/(c**2*M_mol - R*T)
print(f"The calculated amount of degrees of freedom from the experiment is {u.nominal_values(f)[0]:.2f} +- {u.std_devs(f)[0]:.2f}")
```
K2:

```{python,echo = FALSE, comment = NA}
R = 8.314462618
T = ufloat(294.15,1) #Temperature + error in [K]
c = v_sound_k2
M_mol = 0.02896 #[molare masse] [kg/mol]

f = 2*R*T/(c**2*M_mol - R*T)
print(f"The calculated amount of degrees of freedom from the experiment is {u.nominal_values(f)[0]:.2f} +- {u.std_devs(f)[0]:.2f}")
```
K1, argon:

```{python,echo = FALSE, comment = NA}
R = 8.314462618
T = ufloat(294.15,1) #Temperature + error in [K]
c = v_sound_k1_argon
M_mol = 0.04 #[molare masse] [kg/mol]

f = 2*R*T/(c**2*M_mol - R*T)
print(f"The calculated amount of degrees of freedom from the experiment is {u.nominal_values(f)[0]:.2f} +- {u.std_devs(f)[0]:.2f}")
```
K1, co2:

```{python,echo = FALSE, comment = NA}
R = 8.314462618
T = ufloat(294.15,1) #Temperature + error in [K]
c = v_sound_k1_co2
M_mol = 0.044 #[molare masse] [kg/mol]

f = 2*R*T/(c**2*M_mol - R*T)
print(f"The calculated amount of degrees of freedom from the experiment is {u.nominal_values(f)[0]:.2f} +- {u.std_devs(f)[0]:.2f}")
```

In theory, the gases inside K?,... should have ? degrees of freedom. These constist of ...

## Comparison to simulation and theory

...

## Phase diagram

...

## $C_V$, $Z$ and degrees of freedom of a simulated lennard jones gas

...

## Different temperatures at fixed gas density 

...

## Temperatures above the critical point

...

## Determinantion of $C_V$

...