---
title: "Laboritory session 1"
author: "Benedikt Schneider, Carolina Dobras"
date: "\\today"

header-includes:
  - \usepackage{csquotes}
  - \usepackage{titling}
  - \setlength{\droptitle}{-3cm}
output:
  pdf_document:
    
    highlight: tango
    keep_tex: no
    latex_engine: xelatex
    number_sections: yes
    toc: no
    toc_depth: 2
  html_document:
    
    theme: default
    df_print: paged
    toc: no
    toc_depth: '2'
    number_sections: yes
    highlight: tango
    toc_float: yes
    code_folding: hide
link_citation: yes
mainfont: XCharter
lang: de-DE
---

```{r setup, include=FALSE}
library(knitr)
library(Rcpp)
library(reticulate)
use_python("/usr/bin/python3")
knitr::opts_chunk$set(fig.align='center',
                      fig.width=6.0,
                      fig.height=3.6)
```
```{r, include = FALSE}
Sys.setenv("PKG_CXXFLAGS"="-std=c++17")
library(Rcpp)
```

# Laboratory session 1 

The first experiment is using standing waves inside a long tube in order to measure the speed of sound inside the gas the tube is filled with. The tube and 
experiment settings are specified by the following data: 

Tube length: $L = ? \pm (1.5mm?)$. 

input signal: $V_0 = 35mV \pm ?$

Tubes: 

K1: argon or $CO_2$ at $T = T_{room}$

K2: air at $T = T_{room}$

K3: air at $T = 70\,^\circ\mathrm{C}$ = $243.15K$

K4: air at $T = 50\,^\circ\mathrm{C}$ = $223.15$

$T_{room} = ?\,^\circ\mathrm{C}$ = $?K$

We analyzed the properties of the gas in tube K?

# Tasks

## Meaurement of $\nu _k$

We measured the following resonance frequencies $\nu _k$ between $200 Hz$ and $2 kHz$:

K?: $\nu _k = [? \pm ?, ...]$

The other groups meaured:

K?: $\nu _k = [? \pm ?, ...]$

K?: $\nu _k = [? \pm ?, ...]$

K?: $\nu _k = [? \pm ?, ...]$

## How we estimated the error 

...

## The best linear fit 

In order to find the speed of sound in the gas, we want to use the equation $\nu _n = an + b$. However, we do not know which resonance ferquency we are
measuring, so we do not know which n belongs to which measured $\nu _n$. This is why we set the reference of the measured $\nu _n$ to the first one measured. Therefore we get:

$$\nu _n - \nu _{n_0} = \Delta \nu _{\Delta n} = a\Delta n + b - (a(n_0) + b) = a(n-n_0) + b - b = a\Delta n = \frac{c}{2L} \Delta n$$

So the central relation is: $$\Delta \nu _{\Delta n} = \frac{c}{2L} \Delta n$$ 

where $\Delta n$ starts at 0 for the first measurement and for $k$ measurements continues until $k-1$ for the $k^{th}$ resonance frequency. $\Delta \nu _{\Delta n}$ against $\Delta n$ is plotted and
evaluated in the following:

K?:

```{python, echo = FALSE}
from uncertainties import unumpy as u, covariance_matrix as co, ufloat
import numpy as np
from scipy.optimize import curve_fit
import matplotlib.pyplot as plt

nu = u.uarray([1.1,2,3.1], [0.01,0.1,0.1]) #[Messwerte], [sigmas] !
d_n = np.arange(len(nu))
dnu = u.uarray([],[])

for i in range(len(nu)):
    l = nu[i] - nu[0]
    dnu = np.append(dnu, l)

dnu[0] = ufloat(u.nominal_values(dnu)[0],np.mean(u.std_devs(dnu))) #set std dev of the first difference to the mean std dev because otherwise it would be 0

def lin(x, a, b):
    return a*x +b

popt, pcov = curve_fit(lin, d_n, u.nominal_values(dnu), sigma = co(dnu), absolute_sigma = True) 

a = u.uarray([popt[0]], [np.sqrt(pcov[0,0])])
b = popt[1]

plt.figure()
plt.errorbar(d_n,  u.nominal_values(dnu), yerr = u.std_devs(dnu), ecolor = "darkred", capsize = 5, label = "Datapoints", color = "darkblue", fmt = "o")
plt.plot(d_n, u.nominal_values(a)*d_n + b, label = f"Fitted curve, slope = {u.nominal_values(a)[0]:.2f} +- {u.std_devs(a)[0]:.2f}", color = "darkgreen")
plt.grid()
plt.legend()
plt.xlabel(r"$\Delta n$")
plt.ylabel(r"$\Delta \nu _{\Delta n} [Hz]$")
plt.show()
```

Given the slope of the fit curve, the velocity of sound can be calculated:

```{python, echo = FALSE, comment = NA}
L = ufloat(1,0.1)  #Länge, Fehler auf Länge
v_sound = 2*L*a
print(f"speed of sound = {u.nominal_values(v_sound)[0]:.2f} +- {u.std_devs(v_sound)[0]:.2f} m/s")

```
K?:

```{python, echo = FALSE}
from uncertainties import unumpy as u, covariance_matrix as co, ufloat
import numpy as np
from scipy.optimize import curve_fit
import matplotlib.pyplot as plt

nu = u.uarray([1.1,2,3.1], [0.01,0.1,0.1]) #[Messwerte], [sigmas] !
d_n = np.arange(len(nu))
dnu = u.uarray([],[])

for i in range(len(nu)):
    l = nu[i] - nu[0]
    dnu = np.append(dnu, l)

dnu[0] = ufloat(u.nominal_values(dnu)[0],np.mean(u.std_devs(dnu))) #set std dev of the first difference to the mean std dev because otherwise it would be 0

def lin(x, a, b):
    return a*x +b

popt, pcov = curve_fit(lin, d_n, u.nominal_values(dnu), sigma = co(dnu), absolute_sigma = True) 

a = u.uarray([popt[0]], [np.sqrt(pcov[0,0])])
b = popt[1]

plt.figure()
plt.errorbar(d_n,  u.nominal_values(dnu), yerr = u.std_devs(dnu), ecolor = "darkred", capsize = 5, label = "Datapoints", color = "darkblue", fmt = "o")
plt.plot(d_n, u.nominal_values(a)*d_n + b, label = f"Fitted curve, slope = {u.nominal_values(a)[0]:.2f} +- {u.std_devs(a)[0]:.2f}", color = "darkgreen")
plt.grid()
plt.legend()
plt.xlabel(r"$\Delta n$")
plt.ylabel(r"$\Delta \nu _{\Delta n} [Hz]$")
plt.show()
```

Given the slope of the fit curve, the velocity of sound can be calculated:

```{python, echo = FALSE, comment = NA}
L = ufloat(1,0.1)  #Länge, Fehler auf Länge
v_sound = 2*L*a
print(f"speed of sound = {u.nominal_values(v_sound)[0]:.2f} +- {u.std_devs(v_sound)[0]:.2f} m/s")

```
K?:

```{python, echo = FALSE}
from uncertainties import unumpy as u, covariance_matrix as co, ufloat
import numpy as np
from scipy.optimize import curve_fit
import matplotlib.pyplot as plt

nu = u.uarray([1.1,2,3.1], [0.01,0.1,0.1]) #[Messwerte], [sigmas] !
d_n = np.arange(len(nu))
dnu = u.uarray([],[])

for i in range(len(nu)):
    l = nu[i] - nu[0]
    dnu = np.append(dnu, l)

dnu[0] = ufloat(u.nominal_values(dnu)[0],np.mean(u.std_devs(dnu))) #set std dev of the first difference to the mean std dev because otherwise it would be 0

def lin(x, a, b):
    return a*x +b

popt, pcov = curve_fit(lin, d_n, u.nominal_values(dnu), sigma = co(dnu), absolute_sigma = True) 

a = u.uarray([popt[0]], [np.sqrt(pcov[0,0])])
b = popt[1]

plt.figure()
plt.errorbar(d_n,  u.nominal_values(dnu), yerr = u.std_devs(dnu), ecolor = "darkred", capsize = 5, label = "Datapoints", color = "darkblue", fmt = "o")
plt.plot(d_n, u.nominal_values(a)*d_n + b, label = f"Fitted curve, slope = {u.nominal_values(a)[0]:.2f} +- {u.std_devs(a)[0]:.2f}", color = "darkgreen")
plt.grid()
plt.legend()
plt.xlabel(r"$\Delta n$")
plt.ylabel(r"$\Delta \nu _{\Delta n} [Hz]$")
plt.show()
```

Given the slope of the fit curve, the velocity of sound can be calculated:

```{python, echo = FALSE, comment = NA}
L = ufloat(1,0.1)  #Länge, Fehler auf Länge
v_sound = 2*L*a
print(f"speed of sound = {u.nominal_values(v_sound)[0]:.2f} +- {u.std_devs(v_sound)[0]:.2f} m/s")

```
K?:

```{python, echo = FALSE}
from uncertainties import unumpy as u, covariance_matrix as co, ufloat
import numpy as np
from scipy.optimize import curve_fit
import matplotlib.pyplot as plt

nu = u.uarray([1.1,2,3.1], [0.01,0.1,0.1]) #[Messwerte], [sigmas] !
d_n = np.arange(len(nu))
dnu = u.uarray([],[])

for i in range(len(nu)):
    l = nu[i] - nu[0]
    dnu = np.append(dnu, l)

dnu[0] = ufloat(u.nominal_values(dnu)[0],np.mean(u.std_devs(dnu))) #set std dev of the first difference to the mean std dev because otherwise it would be 0

def lin(x, a, b):
    return a*x +b

popt, pcov = curve_fit(lin, d_n, u.nominal_values(dnu), sigma = co(dnu), absolute_sigma = True) 

a = u.uarray([popt[0]], [np.sqrt(pcov[0,0])])
b = popt[1]

plt.figure()
plt.errorbar(d_n,  u.nominal_values(dnu), yerr = u.std_devs(dnu), ecolor = "darkred", capsize = 5, label = "Datapoints", color = "darkblue", fmt = "o")
plt.plot(d_n, u.nominal_values(a)*d_n + b, label = f"Fitted curve, slope = {u.nominal_values(a)[0]:.2f} +- {u.std_devs(a)[0]:.2f}", color = "darkgreen")
plt.grid()
plt.legend()
plt.xlabel(r"$\Delta n$")
plt.ylabel(r"$\Delta \nu _{\Delta n} [Hz]$")
plt.show()
```

Given the slope of the fit curve, the velocity of sound can be calculated:

```{python, echo = FALSE, comment = NA}
L = ufloat(1,0.1)  #Länge, Fehler auf Länge
v_sound = 2*L*a
print(f"speed of sound = {u.nominal_values(v_sound)[0]:.2f} +- {u.std_devs(v_sound)[0]:.2f} m/s")

```
## Most important contribution to uncertainty

...

## Degrees of freedom

For an ideal gas, the velocity of sound is given by $$c(T) = \sqrt{\frac{(f+2)RT}{fM_{mol}}}$$ where $f$ are the degrees of freedom, $R$ is the gas constant,
$T$ the temperature of the gas and $M_mol$ the molar mass. This equation can be soved for $f$: $$f = \frac{2RT}{c^2M_{mol} + RT}$$
We calculate for $f$:

K?:

```{python,echo = FALSE, comment = NA}
R = 8.314462618
T = ufloat(300,1) #Temperature + error in [K]
c = v_sound
M_mol = 20 #[molare masse]

f = 2*R*T/(c**2*M_mol + R*T)
print(f"The calculated amount of degrees of freedom from the experiment is {u.nominal_values(f)[0]:.2f} +- {u.std_devs(f)[0]:.2f}")
```
K?:

```{python,echo = FALSE, comment = NA}
R = 8.314462618
T = ufloat(300,1) #Temperature + error in [K]
c = v_sound
M_mol = 20 #[molare masse]

f = 2*R*T/(c**2*M_mol + R*T)
print(f"The calculated amount of degrees of freedom from the experiment is {u.nominal_values(f)[0]:.2f} +- {u.std_devs(f)[0]:.2f}")
```
K?:

```{python,echo = FALSE, comment = NA}
R = 8.314462618
T = ufloat(300,1) #Temperature + error in [K]
c = v_sound
M_mol = 20 #[molare masse]

f = 2*R*T/(c**2*M_mol + R*T)
print(f"The calculated amount of degrees of freedom from the experiment is {u.nominal_values(f)[0]:.2f} +- {u.std_devs(f)[0]:.2f}")
```
K?:

```{python,echo = FALSE, comment = NA}
R = 8.314462618
T = ufloat(300,1) #Temperature + error in [K]
c = v_sound
M_mol = 20 #[molare masse]

f = 2*R*T/(c**2*M_mol + R*T)
print(f"The calculated amount of degrees of freedom from the experiment is {u.nominal_values(f)[0]:.2f} +- {u.std_devs(f)[0]:.2f}")
```

In theory, the gases inside K?,... should have ? degrees of freedom. These constist of ...

## Comparison to simulation and theory

...

## Phase diagram

...

## $C_V$, $Z$ and degrees of freedom of a simulated lennard jones gas

...

## Different temperatures at fixed gas density 

...

## Temperatures above the critical point

...

## Determinantion of $C_V$

...